---
title: "Reproducible code for bonus rounds"
author: "Eric"
date: "May 16, 2018"
output: html_document
---
###Load packages and data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Let's first load the necessary packages, plyr for data handling, lsr to calclulate Cohen's d, BayesFactor for the BF for a t-test, and jsonlite for reading in the json files.

```{r, results = "hide", message=FALSE}
packages <- c('plyr', 'lsr', 'BayesFactor', 'jsonlite')
lapply(packages, require, character.only = TRUE)
```

Now let's read in the data as well as the json files of the ground truth to calculate the errors.
```{r}
#rough and smooth environments
environments1 <- fromJSON("kernelRough.json")
environments2 <- fromJSON("kernelSmooth.json")

#split up json for rough and smooth enviornments
roughEnvironments <- lapply(environments1, FUN=function(x) matrix(as.numeric(unlist(x)), ncol=3, byrow=TRUE, dimnames=list(seq(1,64), c('x2', 'y', 'x1'))))
smoothEnvironments <- lapply(environments2, FUN=function(x) matrix(as.numeric(unlist(x)), ncol=3, byrow=TRUE, dimnames=list(seq(1,64), c('x2', 'y', 'x1' ))))


#data
myjson<-fromJSON("kwg.json")

#initialize empty frame
dat<-data.frame(id=numeric(), age=numeric(), x=numeric(), y=numeric(), est=numeric(), 
                cer=numeric(), cond=numeric(), truth=numeric(), chosen=numeric())

#loop through json
for (i in 1:myjson$count){
  #x-y-locations
  x<-as.numeric(myjson$records$data$bonusLevel$bonusCells[[i]]$x)
  y<-as.numeric(myjson$records$data$bonusLevel$bonusCells[[i]]$y)
  #estimates
  est<-as.numeric(myjson$records$data$bonusLevel$bonusCells[[i]]$givenValue)
  #uncertainties
  cert<-as.numeric(myjson$records$data$bonusLevel$bonusCells[[i]]$howSecure)
  #environments
  env<-myjson$records$data$envOrder[[i]][10]
  #condition
  condition<-myjson$records$data$condition[i]
  #get smooth
  if(condition==0){ee<-data.frame(smoothEnvironments[[env+1]])}
  if(condition==1){ee<-data.frame(roughEnvironments[[env+1]])}
  #initialize truth
  truth<-rep(0, length(x))
  #loop over truth
  for (k in seq_along(truth)){
    #get truth
    truth[k]<-ee$y[x[k]==ee$x1 & y[k]==ee$x2]
  }
  #truth rescaled
  truth<-truth*myjson$records$data$scale[[i]][10]+5
  #initialize chosen
  chosen<-rep(0, 5)
  #which one did they choose
  ch<-myjson$records$data$bonusLevel$finalChosenCell[i,]
  #mark chosen one by a 1
  chosen[x==ch$x & y ==ch$y]<-1
  #dummy frame
  dummy<-data.frame(id=rep(i, 5), age=rep(myjson$records$data$age[i], 5), x=x, y=y,
                    est=est, cer=cert, cond=rep(condition, 5), truth=truth, chosen=chosen)
  #bind data frames together
  dat<-rbind(dat, dummy)
}

#first 10 was us!
dat<-subset(dat, id>10)
#age has to be at least 7
dat<-subset(dat, age>=7)

#get absolute error
dat$error<-abs(dat$est-dat$truth)

#create age group
dat$agegroup<-ifelse(dat$age<9, "7-8", dat$age)
dat$agegroup<-ifelse(dat$age>=9 & dat$age <12, "9-11", dat$agegroup)
dat$agegroup<-ifelse(dat$age>18, ">18", dat$agegroup)

#Condition
dat$cond<-ifelse(dat$cond==1, "Rough", "Smooth")
```


###Error of predictions
A first thing we want to check, is if the error depends on the condition, i.e. do participants show larger errors in the rough condition?

```{r}
#Error by condition
de<-ddply(dat, ~id+cond, summarize, mu=mean(error))
#Rough vs. smooth
t.test(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu, var.equal = TRUE)
cohensD(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu)
ttestBF(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu)
```
As expected, the error for the rough condition is higher than for the smooth. This means that more correlation in the environment helped participants to make better predictions.

Next, we want to check if the different age groups differ in ther prediction errors.
```{r}
#Error by age
dage<-ddply(dat, ~id+agegroup,summarize, mu=mean(error))

#adults vs. old children
t.test(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu, var.equal = TRUE)
cohensD(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu)
ttestBF(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu)

#old children vs. young children
t.test(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu, var.equal = TRUE)
cohensD(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu)
ttestBF(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu)

```

Interestingly, older children generated better predictions than adults and also better prediction than younger children.


###Certainty judgements
Next, we are interested in participants' certainty estimates.

First, do the certainty estimates differ by condition?

```{r}
#Certainty judgements
de<-ddply(dat, ~id+cond, summarize, mu=mean(cer))
#Smooth vs. Rough
t.test(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu, var.equal = TRUE)
cohensD(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu)
ttestBF(subset(de, cond=="Rough")$mu, subset(de, cond=="Smooth")$mu)
```

Turns out they don't, participants were equally certain about their predictions in both conditions.

Did certainty estimates differ by age?
```{r}
#certainty by age group
dage<-ddply(dat, ~id+agegroup,summarize, mu=mean(cer))
#make age a factor
dage$agegroup<-as.factor(dage$agegroup)
#ANOVA Bayes Factor
anovaBF(mu~agegroup, data=dage)

```
Also no diffference.

Finally, we are interested in participants' choices, do they differ somehow? To asses this, we divide the values of the chosen options by the sum of all option for both the predictions and the certainty judgements separately.
```{r}
#standardized characeristics of chosen options
ds<-ddply(dat, ~id+agegroup, summarize, cer=cer[chosen==1]/sum(cer), mu=est[chosen==1]/sum(est))
ds$agegroup<-factor(ds$agegroup)

#difference in predicted value
anovaBF(mu~agegroup, data=ds)

#difference in uncertainty
ds$cer<-ifelse(is.na(ds$ce), 0.2, ds$ce)

#overall
anovaBF(cer~agegroup, data=ds)

#between the two children groups
t.test(subset(ds, agegroup=="9-11")$cer, subset(ds, agegroup=="7-8")$cer, var.equal = TRUE)
cohensD(subset(ds, agegroup=="9-11")$cer, subset(ds, agegroup=="7-8")$cer)
ttestBF(subset(ds, agegroup=="9-11")$cer, subset(ds, agegroup=="7-8")$cer)

#between the adults and young kids
t.test(subset(ds, agegroup==">18")$cer, subset(ds, agegroup=="7-8")$cer, var.equal = TRUE)
cohensD(subset(ds, agegroup==">18")$cer, subset(ds, agegroup=="7-8")$cer)
ttestBF(subset(ds, agegroup==">18")$cer, subset(ds, agegroup=="7-8")$cer)

```
This shows that participants did not differ in their standardized predictions, i.e. different age groups did not choose higher values on average. However, there's is somewhat of a difference in certainty, where older children and adults differ marginally and younger kids and adults differ substantially.