---
title: "Reproducible code for behavioral results"
author: "Eric Schulz"
date: "May 15, 2018"
output: html_document
---
###Load packages and take a first look at the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Let's first load the necessary packages, plyr for data handling, lsr to calclulate Cohen's d, and BayesFactor for the BF for a t-test.

```{r, results = "hide", message=FALSE}
packages <- c('plyr', 'lsr', 'BayesFactor')
lapply(packages, require, character.only = TRUE)
```


Now let's load the data and have a look at the structure:
```{r}
#read in data
dat<-read.csv("kwgdata.csv")
#show first 5 entries
head(dat, 5)
```
In this frame, id is the participant's id, cond is the condition (i.e. the correlational structure of the environments that participants experienced), age is the raw age, x and y are the sampled coordinates on the grid, z is the outcome (reward), trial is the trial number, round is the round number, and ageroup is one of the three agegroups, 7-8, 9-11, and >18.


###Performance by condition
A first thing to test is the influence of condition on performance.
```{r}
#get mean by condition
dcond<-ddply(dat, ~id+cond,summarize, mu=mean(z))
#t-test between rough and smooth
t.test(subset(dcond, cond=="Rough")$mu, subset(dcond, cond=="Smooth")$mu, var.equal = TRUE)
#Effect size
cohensD(subset(dcond, cond=="Rough")$mu, subset(dcond, cond=="Smooth")$mu)
#Bayes factor
ttestBF(subset(dcond, cond=="Rough")$mu, subset(dcond, cond=="Smooth")$mu)
```

This inidcates a strong effect of condition onto performance. As expected, smooth environments lead to better performance than rough environments.

###Performance by age group
Next, we are going to assess if the different age groups performed differently.
```{r}
#test age, exclude trial 0 cause that's same for all.
dage<-ddply(subset(dat,trial>0), ~id+agegroup,summarize, mu=mean(z))

#adults vs. older children
##t-test
t.test(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu, var.equal = TRUE)
##effect size
cohensD(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu)
##Bayes factor
ttestBF(subset(dage, agegroup==">18")$mu, subset(dage, agegroup=="9-11")$mu)

#older children vs. adults
##t-test
t.test(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu, var.equal = TRUE)
##effect size
cohensD(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu)
##Bayes factor
ttestBF(subset(dage, agegroup=="9-11")$mu, subset(dage, agegroup=="7-8")$mu)

```
This shows that adults performed better than older children, who in turn performed somewhat better than younger children.


###Distance of sampling behavior
As we're interested in exploration-exploitation, let's check out the distances between participants' consecutive samples.

```{r}
#Initialize distance to be 0
dat$dist<-0
#loop th
for (i in 2:nrow(dat)){
  #absolute difference between current and previous move
  dat$dist[i]<-abs(dat$x[i]-dat$x[i-1])+abs(dat$y[i]-dat$y[i-1])
  #we can only do this, if we then alway exclude trial 0.
}
```

Let's check out how this distance differs between conditions.
```{r}
#Distance by condition
#distance by condition and id
ddist<-ddply(subset(dat,trial>0), ~id+cond,summarize, mu=mean(dist))
#t-test between condition's distances
t.test(subset(ddist, cond=="Rough")$mu, subset(ddist, cond=="Smooth")$mu, var.equal = TRUE)
#effect size
cohensD(subset(ddist, cond=="Rough")$mu, subset(ddist, cond=="Smooth")$mu)
#Bayes factor
ttestBF(subset(ddist, cond=="Rough")$mu, subset(ddist, cond=="Smooth")$mu)
```
We can see that participants in the smooth condition sampled more locally (smaller average distance) than participants in the rough condition.

Let's check out the different age groups now.
```{r}
#Distance by age group
ddista<-ddply(subset(dat,trial>0), ~id+agegroup,summarize, mu=mean(dist))

#adults vs. older children
##t-test
t.test(subset(ddista, agegroup==">18")$mu, subset(ddista, agegroup=="9-11")$mu, var.equal = TRUE)
##effect size
cohensD(subset(ddista, agegroup==">18")$mu, subset(ddista, agegroup=="9-11")$mu)
##Bayes factor
ttestBF(subset(ddista, agegroup==">18")$mu, subset(ddista, agegroup=="9-11")$mu)

#older vs. younger children
##t-test
t.test(subset(ddista, agegroup=="9-11")$mu, subset(ddista, agegroup=="7-8")$mu, var.equal = TRUE)
##effect size
cohensD(subset(ddista, agegroup=="9-11")$mu, subset(ddista, agegroup=="7-8")$mu)
##Bayes factor
ttestBF(subset(ddista, agegroup=="9-11")$mu, subset(ddista, agegroup=="7-8")$mu)

```
So adults sampled more locally than older children, but older children did not sample more locally than younger children.

###Number of unique options
Another inidicator of exploration is how many unique options participants sampled overall. So let's first create a variable that measures that.

```{r}
#indicator sting for tile position
dat$comb<-paste(dat$x, dat$y)
#number of unique options sampled per round
dnopt<-ddply(dat, ~id+agegroup+round, summarize, l=length(unique(comb)))
#summarize per id with age group
dnopt<-ddply(dnopt, ~id+agegroup, summarize, mu=mean(l))
```

This just measures how many unique options every participants sampled in total. So let's see how this variable differs between the different age groups.
```{r}
#adults vs. older children
##t-test
t.test(subset(dnopt, agegroup==">18")$mu, subset(dnopt, agegroup=="9-11")$mu, var.equal = TRUE)
##effect size
cohensD(subset(dnopt, agegroup==">18")$mu, subset(dnopt, agegroup=="9-11")$mu)
##Bayes factor
ttestBF(subset(dnopt, agegroup==">18")$mu, subset(dnopt, agegroup=="9-11")$mu)

#older vs. younger children
##t-test
t.test(subset(dnopt, agegroup=="9-11")$mu, subset(dnopt, agegroup=="7-8")$mu, var.equal = TRUE)
##effect size
cohensD(subset(dnopt, agegroup=="9-11")$mu, subset(dnopt, agegroup=="7-8")$mu)
##Bayes factor
ttestBF(subset(dnopt, agegroup=="9-11")$mu, subset(dnopt, agegroup=="7-8")$mu)

```

###Learning curves
Finally, we want to assess learning curves, i.e. how much participants got better over trials.

```{r}

#Learning curves assessed by Spearman's rho
dt<-ddply(dat, ~id+cond+agegroup, summarize, c=cor(trial, z, method="spearman"))
```

We calculate Spearman's rho, i.e. a rank correlation, between trials and rewards for each participant. Although not perfect, this can account for the non-linearity of learning curves.

Let's first check if this correlation is different from 0 at all.
```{r}
#Learning curves assessed by Spearman's rho
#different from 0?
t.test(dt$c)
ttestBF(dt$c)
```
Clearly, participants are getting better over trials as exepcted.

Let's check for differences between conditions next.
```{r}
#difference between conditions
##t-test
t.test(subset(dt, cond=="Rough")$c, subset(dt, cond=="Smooth")$c, var.equal = TRUE)
##effect size
cohensD(subset(dt, cond=="Rough")$c, subset(dt, cond=="Smooth")$c)
##Bayes factor
ttestBF(subset(dt, cond=="Rough")$c, subset(dt, cond=="Smooth")$c)
```
There was no difference in learning speed between the different conditions. 

Finally, let's compar the different age groups.
```{r}
#Adults vs. older children
##t-test
t.test(subset(dt, agegroup==">18")$c, subset(dt, agegroup=="9-11")$c, var.equal = TRUE)
##effect size
cohensD(subset(dt, agegroup==">18")$c, subset(dt, agegroup=="9-11")$c)
##Bayes factor
ttestBF(subset(dt, agegroup==">18")$c, subset(dt, agegroup=="9-11")$c)

#Older children vs. younger children
##t-test
t.test(subset(dt, agegroup=="7-8")$c, subset(dt, agegroup=="9-11")$c, var.equal = TRUE)
##effect size
cohensD(subset(dt, agegroup=="7-8")$c, subset(dt, agegroup=="9-11")$c)
##Bayes factor
ttestBF(subset(dt, agegroup=="7-8")$c, subset(dt, agegroup=="9-11")$c)
```
So this correlation is higher for adults than for older children, while there is no difference between older and younger children.